\documentclass{article}
\usepackage{amsmath,amssymb}
\usepackage{color}
\usepackage{bm}
\usepackage[sort&compress]{natbib}
\usepackage[normalem]{ulem}	% Part of the standard distribution


\begin{document}
\textbf{Question:}
\begin{quotation}
The main quantity that the paper focuses on is the thickness of the film between the bubbles and the
wall. As shown by Figure 8 and as mentioned by the Authors in Section 4, the film thickness varies
along the bubble. However, the Authors do not explain the choice of the location of the measurement
of the film thickness. This choice is of primary importance when comparing results with those of
the
literature and in particular with works on finger propagation in which the film thickness is
measured
”at infinity”.
\end{quotation}

\textbf{Answer:} The following section is added to the text 
\begin{quotation}
We chose to measure the film thickness in the middle of the bubble. The
classical Bretherton formulation is for the measurement of the film thickness at ``infinity''.
However, \citet{cerro-bubble-train} indicate that for the gravity-driven bubble train flow in tubes
it is enough for bubble to have a length of two-three diameters to apply analysis for very long
bubbles. In comparison with the axisymmetric circular geometry, for three-dimensional square shaped
microchannels \citet{heil-threedim} indicate the measurement of the film thickness at $5.5$
channel heights from the front bubble tip. This distance was shown sufficient for the
film thickness to establish itself for the capillary number range $Ca<4$. We chose to measure
the film thickness in the middle of the bubble, which is located at least at the distance of $2.5-3$
channel heights from the front bubble tip.  Moreover, the film thickness examination, see Section
\ref{section:film:thickness:variation}, shows that for the capillary range of interest, i.e.
$0.08\leq Ca \leq 1$, the standard deviation from the bubble middle film thickness is around $7$
 percent for small capillary numbers ($Ca=0.05$) and less than $1$ percent for larger ones
($Ca \geq 0.1$) at the distances ranging from one channel height from the rear tip upto one channel
height to the front tip. This certainly validates the point of choice to measure bubble thickness.
\end{quotation}
 
\textbf{Question:}
\begin{quotation}
It may be appropriate to introduce Figure 8 on thickness variation before the main results on ﬁlm
thickness.
\end{quotation}

\textbf{Answer:} We agree and moved the section before the results on the capillary number
dependency.


\textbf{Question:}
\begin{quotation}
The lines 54-55 indicating that ”numerical simulations and experimental studies showed [...]
Reynolds
number effects on the ﬁlm thickness for capillary numbers larger than 0.003” is in contradiction
with
the lines 32-33 which state ”negligible Reynolds number effects on the film thickness for a
relatively
wide range of Reynolds numbers.”
\end{quotation}

\textbf{Answer:} The following paragraph is added:
\begin{quotation}
Moreover, the results of \citet{giavedoni-numerical} and \citet{heil-bretherton} show
{\color{red} insignificant} Reynolds number effects on the film thickness for a relatively wide
range of Reynolds
numbers. 
{\color{red} For example, \citet{giavedoni-numerical} suggested that the Reynolds number effects
are negligible for $Ca\leq0.05$ and have moderate changes for $Ca>0.05$ in the range of Reynolds
number from $0$ to $70$. Later on, citet{heil-bretherton} extended results upto Reynolds number
$300$. The indicate that while the influence of the established film thickness is insignificant
($7$ percent from the film thickness measured at $Re=0$). However, the Reynolds number influence
significantly the pressure distrubution and the flow field near the front bubble tip. In
the present simulations the largest Reynolds number we achieve is less than $20$ and it allows to
neglect the
inertia effects.} 

\citet{kreutzer-pressure-drop} showed a deviation from the $Ca^{2/3}$ rule
{\color{red}\sout{and Reynolds number 
effects on the film thickness}} for capillary numbers larger than $0.003$.
\end{quotation}

\textbf{Question:}
\begin{quotation}
Although inertia has been shown to have a minor effect on ﬁlm thickness, inertia does influence
some flow variables in the Bretherton’s problem (see Ref. 3 for instance). To check whether inertia
is important or not, and to discuss the assumption of uniform density, the Authors should provide
the reader with the Reynolds numbers associated with each simulation. This is also required for the
consistency of Figure 7 in which results are compared with other works of the literature.
\end{quotation}

\textbf{Answer:} Thank you for pointing it out. We included the discussions of the Reynolds
effects in the answer to the question above. As well, we added Reynolds numbers to the Table for
the capillary range regime and the grid depedency study. 

\textbf{Question:}
\begin{quotation}
In this work, the diphasic flow is driven by a body force. The body force is therefore the control
parameter in the simulations. The Authors should give more detail about this body force. Indeed,
although the estimation of the body force from the pressure gradient is well described, it is not
clear
where this body force appears in the governing equations (Equations 2, 3 and 4). Does this body
force correspond to the variable $F_i$ in Equation 2? Should not this force or its equivalent appear
in
the Navier-Stokes equations (Equation 4)?
\end{quotation}

\textbf{Answer:} We updated equation (4) to include the force contribution. As well, the definition
of $P_{\alpha\beta}$ is added. $F_i$ is the force population which corresponds to the force
inclusion $\bm{F}$. We added the citation which describes how the force population is formed and
mimics the force behavior.

\textbf{Question:}
\begin{quotation}
I would suggest to calculate the Bond number to check if the shapes of the bubbles and the main
results of the paper are not inﬂuenced.
\end{quotation}

\textbf{Answer:} 
The maximum Bond number $Bo=\frac{\rho g N_y^2}{\gamma}$ in the
current simulation equals to $7$, which is less than the number when the shape of the bubble is
affected \cite{zheng-large-ratio}. However, the point is that the equation of state in the lattice
Boltzmann framework is connected with density. Thus usually either pressure (i.e. density) boundary
conditions are prescribed or the body force. If they are prescribed simultaneously then
we can consider the influence of the Bond number on the simulations. Otherwise prescribing just the
body force is analogous to prescribing the pressure gradient $\frac{\mathrm{d}P}{\mathrm{d}x}$ in
the Navier-Stokes equation. Thus, it's not possbile the Bond number effect. It is because of the
coupled nature of the pressure and density in the LBM framework. 

\textbf{Question:}
\begin{quotation}
The use of the variable $Ca_{lit}$ and the calculation of the corresponding pressure gradient yield
estima
tions of the body force to apply and of the ﬁlm thickness to choose grid resolution. The
presentation
of the corresponding procedure is repeated twice in Section 5-1 for $Ca=0.005$ and $Ca=0.05$. It is
again described in a diﬀerent way in Section 5-4. Similar equations for the calculation of the
pressure
gradient are thus given in Equations 6, 14 and 16. The Authors should summarize these equations
into one unless they explain their differences. In fact, it would probably be straightforward not to
detail the calculation of estimated variables since they are intermediary variables which allow the
design of the benchmark but do not inﬂuence directly the results. Also, the Authors could possibly
say they control and impose directly the body force without describing all the estimation procedure.
\end{quotation}

\textbf{Answer:} Thank you for the suggestion. We agree that there is redundant information
about initialization and this was taken into the account by reducing repetitive text and
introducing step-by-step explations, see Section about initialization. We also added the
recommendations for controlling parameters in the same section.  The
dimensionalization of the problem and proper benchmarking is quite difficult in the lattice
Boltzmann framework. You can refer to the lbmethod.org/forum where half of all questions related to
the lattice Boltzmann are refered to the non-dimensionalization procedure. Therefore, we decided to
keep suggestions for controlling and initialization of simulations as informative guides for the
LBM community.

\textbf{Question:}
\begin{quotation}
 Moreover, the estimations of the pressure gradient for the initialisation of the scheme are
obtained
under the assumption of an axisymmetric Poiseuille ﬂow. However, among the ”Key words”, one can
read the key word ”Flow between plates” in agreement with the 2D non axisymmetric simulations
that are done in this work. The Authors should discuss this choice of estimating the pressure
gradient with the axisymmetric assumption instead of a Hele-Shaw flow (factor 1/12 instead of 1/8
in Equation 5).
\end{quotation}

\textbf{Answer:} While the Hele-Shaw flow implies the flow with the infinitisemal distance between
plates, in our situation the size of the bubble can be compared with the channel height ($5$ times
larger) and that's why the assumption of the infinitisemal distance between plates is too strict.
Moreover, after recalculating results with the velocity of the interface the assumption of the
Poiseulle flow is valid and shows good accuracy.

\textbf{Question:}
\begin{quotation}
The Section on LBM is concise and this seems a reasonable choice when publishing such a work in a
journal of applied science. The Authors would improve signiﬁcantly the manuscript by deﬁning more
rigorously some of the variables of the model. For instance, the Authors refer to surface tension in
lines 106, 111 and 123 via the parameter k, the weights w and the quantity 8kA/9 respectively.
A suggestion is to emphasize within an array the links between physical parameters and the model
eq
parameters. The variables i, fieq , gi should be deﬁned and the Equations for f and g in Equation
3 should be described (which equation corresponds to continuity equation...). Equation 2 could
perhaps be rewritten with the use of a generic variable in order to remove any ambiguity with the
physical variable f of Equation 3. The notation i = 1 ÷ 8 is unusual, it would be clearer to use a
classical notation such as 1 ≤ i ≤ 8. The sound speed c2 should be deﬁned just after its
introduction in line 107.
\end{quotation}

\textbf{Answer:} We wrote an additional paragraph emphasizing the connection of the physical world
parameters with the lattice Botlzmann parameters. As well more explanations are put towards the
interface surface tension description and inclusion it to the model. The reference to the
continuity and the Navier-Stokes equations are added. The equation (2) is not rewritten in terms of
generic variables as far as it's a historical LBM notation for the Navier-Stokes and phase
governing equations. The notations are corrected.

\textbf{Question:}
\begin{quotation}
The name of the Section ”Lattice Boltzmann benchmark” lets the reader think that this Section
would describe the numerical procedure (assumptions, choice of the geometry and the grid, boundary
conditions, control parameters, initialization of the scheme). This is not the case because this
Section
develops the challenges and the difficulties of using the Lattice Boltzmann method to bubble ﬂows.
This Section should be renamed or reorganized.
\end{quotation}

\textbf{Answer:} We renamed the section to ``Approach to benchmark''.

\textbf{Question:}
\begin{quotation}
The Authors should make a distinction between variables in lattice units and dimensional or non
dimensional variables. For instance, in line 195, the definition $H_{eff} = N_y − 2$ where $N_y$ is
the
number of nodes in the vertical direction indicates that $H_{eff}$ has lattice units. However, this
is in
contradiction with the use of $H_{eff}$ in Equation 5. The Authors should deﬁne each variable
without
any ambiguity on its units. Besides, the unit of $U_{bubble}$ should also be given. These
improvements
would ﬁt well in the Subsection named ”The nondimensionalization and initialization procedure”
which does not in fact introduce much nondimensionalisation procedure.
\end{quotation}

\textbf{Answer:} We added a decription paragraph in the Subsection ``The nondimensionalization and
initialization procedure'' about the approach to units. In fact, all equations are the same in the
lattice Boltzmann framework or in a physical framework. One can substitute any quantity in
preferred units as soon as capillary number is matched in the LBM and the physical world. 

\textbf{Question:}
\begin{quotation}
Some notations are unclear. For instance, the ﬁlm thickness is denoted as h∞ in Equation 7 whereas
it is denoted δ in most of the paper. The notations w or δsim are also found before line 381 or in
Table 3 respectively. The thickness of the interface is denoted ξ in Figure 1 but the same variable
is
denoted 5ξ in Table 1. The speed of the bubble is denoted Ububble in some places and Ub in others.
The variable Uin Equation 5 does not seem to be deﬁned.
\end{quotation}

\texbf{Answer:} We thoroughly corrected all the inconsistencies.

\textbf{Question:}
\begin{quotation}
Three ﬁgures in the manuscript are scanned from the litterature (Ref. 3, 10 and 16). This is
unusual
for an article which is not a review. I encourage the Authors to remove at least the two ﬁrst ones
which are not essential.
\end{quotation}

\textbf{Answer:} We eliminated necessary pictures and references to them.

\textbf{Question:} 
\begin{quotation}
In Figure 7, the points of data reported from the numerical work of Ref. 3 exhibit a surprisingly
noisy trend that corresponds more to an experimental study. The Authors should explain how they
obtain this data and plot them.
\end{quotation}

\textbf{Answer:} We added a description to the figure caption explaining data acquisition.

\textbf{Question:} 
\begin{quotation}
The plots of Figure 8 are of less quality than the other ﬁgures of the paper.
\end{quotation}

\textbf{Answer:} Unfortunately, it was not possible to extract data as a picture from the paper by
\citet{sehgal-microchannel}. Thus, the curves were extracted from the graphs and represented in a
new figure scaled to the present simulations for comparison reasons.

\textbf{Question:}
\begin{quotation}
The work of Yang seems to be close from the present study since it involves the
solution of Bretherton
ﬂow with a Lattice Boltzmann method. The Authors should discuss this paper and emphasize the
differences between their work and this paper.
\end{quotation}

\textbf{Answer:} We outlined the differences in the introduction section.

\textbf{Question:} It would be interesting if the Authors could mention a study in which a level-set
method is used in
order to complete the bibliography.

\textbf{Answer:} We added references to the introduction section.

\textbf{Questions:} 
\begin{quotation}
A secondary quantity of interest is the bubble speed or in a non-dimensional
form the capillary
number. This quantity is measured at the center of the bubbles. The Authors should describe how
this quantity as well as the center of the bubble are measured.
\end{quotation}

\textbf{Answer:} Thank you for pointing that. The classical Bretherton problem
$U_{\mathrm{bubble}}$ is the bubble velocity, i.e. the interface velocity. We recalculate all the
results taking into the consideration the interface velocity at the bubble tip. Qualitatively, the
results are the same. It was found that Poiseuille profile assumption is improved significantly
after recalculation.

\textbf{Question:} 
\begin{quotation}
 If possible, the Authors should present results showing that steady state is reached. It would
also
be interesting if the Authors show some patterns of the ﬂow by plotting for instance the vector ﬁeld
of velocity around the bubbles.
\end{quotation}

\textbf{Answer:} We added additional section which shows the change of flow pattern. 

\textbf{Question:}
\begin{quotation}
In the Abstract, the Authors should give the range of real capillary numbers they measured instead
of the targeted values (Calit ).
\end{quotation}

\textbf{Answer:} We changed it to the measured capillary numbers.

\textbf{Question:} 
\begin{quotation}
The Authors should discuss more their results (Sections 5-4, 5-5, 5-6). Did the Authors measure the
pressure drops at the front and rear interfaces of the bubbles?
\end{quotation}

 

\bibliographystyle{unsrtnat}
\bibliography{answers}

\end{document}
